### 2020-01-30 wuhan.tex

Here, we show that a wide range of approaches to estimating \Ro\ can be understood and compared in terms of estimates for three quantities: the exponential growth rate $r$, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$.
The generation interval, which is defined as the time between when an individual becomes infected and when that individual infects another individual \citep{svensson2007note}, plays a key role in shaping the relationship between $r$ and \Ro\ \citep{wearing2005appropriate, roberts2007model, wallinga2007generation, park2019practical};
early in an epidemic, information is scarce and, inevitably, there is large uncertainty around contact tracing (affecting the estimates of generation intervals) and case reports (affecting the estimates of exponential growth rate).
We suggest that disease modelers should make sure their assumptions about these three quantities are clear and reasonable, and that estimates of uncertainty should propagate error from all three sources.

@@

For each study, we construct a family of parameter sets by drawing 100,000 random samples from the probability distributions (Table 2) that represent the estimates of \Ro and the assumed values of $\bar G$ and $\kappa$ and calculating the exponential growth rate $r$ via the inverse of \eref{gamma}:
\begin{equation}
r_i = \frac{{\mathcal R_{0i}}^{\kappa_i} - 1}{\kappa_i \bar{G}_i}.
\end{equation}
This allows us to approximate the probability distributions of the estimated exponential growth rates by each study.

\dc{I think I understand why you do that. But I’m worried the uncertainty for r comes
artificially from whatever methods the Studies used, not the actual observation process we
would have by observing (imperfectly) directly the cases. I guess the studies have all more or
less the same incidence data... why not estimating uncertainty for r directly from incidence}

@@

\bmb{general comments (Daniel may have taken notes on these already):
  \begin{itemize}
  \item (\textbf{later on}) try flatter-at-origin priors (e.g. half-Cauchy)? Stan implementation?
  \item comment on broader uncertainty context (King/Rohani on observation + process error, major reporting irregularities, non-stationarity \ldots)
  \item cite \cite{elderd2006uncertainty} on point estimates
  \end{itemize}
}

>>>>

Here, we show that a wide range of approaches to estimating \Ro\ can be understood and compared in terms of estimates for three quantities: the exponential growth rate $r$, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$.
The generation interval, which is defined as the time between when an individual becomes infected and when that individual infects another individual \citep{svensson2007note}, plays a key role in shaping the relationship between $r$ and \Ro\ \citep{wearing2005appropriate, roberts2007model, wallinga2007generation, park2019practical};
early in an epidemic, information is scarce and, inevitably, there is large uncertainty around case reports (affecting the estimates of exponential growth rate) and contact tracing (affecting the estimates of generation intervals).
We suggest that disease modelers should make sure their assumptions about these three quantities are clear and reasonable, and that estimates of uncertainty should propagate error from all three sources.

@@

\subsection{Description of the studies}

We reviewed six modeling studies of the novel coronavirus outbreak that were published online between January 24th, 2020 and January 26th, 2020 (Table 1).
Five studies \citep{liuncov, majumderncov, readncov, riouncov, zhaoncov} were upload on preprint servers (bioRxiv, medRxiv, and SSRN), and one report was published from Imperial College London \citep{imaincov}.
\cite{imaincov} and \cite{rioucov} estimated the basic reproduction number by simulating branching process models

majumder 26th
read 21st
zhao 22nd
liu 23rd
Two based on 18th.

\subsection{Statistical framework}

For each study, we construct a family of parameter sets by drawing 100,000 random samples from the probability distributions (Table 2) that represent the estimates of \Ro and the assumed values of $\bar G$ and $\kappa$ and calculating the exponential growth rate $r$ via the inverse of \eref{gamma}:
\begin{equation}
r_i = \frac{{\mathcal R_{0i}}^{\kappa_i} - 1}{\kappa_i \bar{G}_i}.
\end{equation}
This allows us to approximate the probability distributions of the estimated exponential growth rates by each study;
uncertainties in the probability distributions that we calculate for the estimated exponential growth rates will reflect the methods and assumptions that the studies rely on.

\dc{I think I understand why you do that. But I’m worried the uncertainty for r comes
artificially from whatever methods the Studies used, not the actual observation process we
would have by observing (imperfectly) directly the cases. I guess the studies have all more or
less the same incidence data... why not estimating uncertainty for r directly from incidence}
\swp{that is exactly the point... we want to be able to reflect their methods... added a sentence about it.}

@@

\bmb{general comments (Daniel may have taken notes on these already):
  \begin{itemize}
  \item (\textbf{later on}) try flatter-at-origin priors (e.g. half-Cauchy)? Stan implementation?
  \item comment on broader uncertainty context (King/Rohani on observation + process error, major reporting irregularities, non-stationarity \ldots)
  \end{itemize}
}
